---
title: "HCD Simulations Write Up"
author: "Audrey Fu Lab"
date: "2024-03-07"
output:
 pdf_document:
   keep_tex: true
   number_sections: true
always_allow_html: true
header-includes:
  - \usepackage{float}
  - \usepackage{multirow}
  - \usepackage{lastpage}
  - \usepackage{fancyhdr}
  # - \pagenumbering{gobble}
  - \pagestyle{fancy}
bibliography: C:/Users/Bruin/Desktop/Research Assistantship/Thesis Proposal Defense/proposal_references.bib
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_chunk$set(fig.width=12, fig.height=10, fig.pos = "H")
#knitr::opts_chunk$set(list(echo = FALSE, eval = FALSE))
library(kableExtra)
library(knitr)
library(latex2exp)
library(gridExtra)
library(ggpubr)
library(ggthemes)
library(plyr)
basepath = 'C:/Users/Bruin/Documents/GitHub/HGRN_repo/Simulated Hierarchies/'

library(ggpubr)
tab = read.csv(paste0(basepath, 'DATA/Toy_examples/Intermediate_examples/Results/MASTER_results.csv'))




```

\section*{Data Simulation}

We simulate hierarchical networks in a top-down approach. We consider several parameters of simulation such as sparsity, noise, and the architecture of the super level graph(s), namely small-world, and scale-free networks [@watts1998collective; @barabasi2003scale].  We simplify our simulations by focusing on basic hierarchies with just one or two hierarchical layers. 
	
In each hierarchy, we start by simulating the top-level nodes in topological order, using either a small world or scale-free network structure [@watts1998collective; @barabasi2003scale]. We define origin nodes as nodes in the topological graph that have no parental input. All origin nodes are simulated from a standard normal distribtion. After generating an initial graph corresponding to the top most layer of the hierarchy, we simulate the middle and bottom layers of the hierarchy by creating groups of new nodes for each parent super-node in the upper level(s). 
	
Each hierarchical level contains community structure nested in the previous layer. The number of offspring nodes generated for each parent node in the level above is chosen from a uniform $\text{unif}(a,b)$ distribution. We also control the connection probabilities both within and between the communities of each hierarchical layer. Once a hierarchical graph is simulated we use the hierarchy to generate the node-feature matrix which represents the expression of $N$ genes in $p$ samples. The number $N$ represents the number of nodes in the observed (bottom) layer of the hierarchy and ranges between $a^{\ell+1}<N<a\times b^\ell$ where $\ell$ represents the number of hierarchical layers. 
	
We consider three sets of hierarchical networks which represent varying difficulty levels for inference:
	
\begin{itemize}
  \item[1.] Complex networks - - used for final simulation assessment 
  \item[2.] Intermediate networks - used for investigative model tuning and performance assessment 
  \item[3.] Simple networks - used for code implementation and debugging
\end{itemize}

\section*{Application to Intermediate Networks}

A summary of the intermediate networks can be found in \textbf{Table 1}. These

\section*{Preliminary Findings}


\newpage
\section*{Figures}
```{r fig1, echo=F, message=F, warning=F, fig.height=8, fig.width=10}

longform = cbind.data.frame(metric = as.factor(c(rep('homogeneity', dim(tab)[1]),
                              rep('completeness', dim(tab)[1]),
                              rep('NMI', dim(tab)[1]))),
                            lambda_top = as.factor(rep(tab$Lambda_top, 3)),
                            lambda_middle = as.factor(rep(tab$Lambda_middle, 3)),
                            network = as.factor(rep(tab$Network_type, 3)),
                            input_graph = as.factor(rep(tab$input_graph, 3)),
                            gamma = as.factor(rep(tab$Gamma, 3)),
                            TOAL_loss = as.factor(rep(tab$TOAL_Loss, 3)),
                            resolution_top = as.factor(rep(tab$Resolution_top, 3)),
                            resolution_middle = as.factor(rep(tab$Resolution_middle, 3)),
                            Top.stats = c(tab$Top_homogeneity, tab$Top_completeness, tab$Top_NMI),
                            Mid.stats = c(tab$Middle_homogeneity, tab$Middle_Completeness, tab$Middle_NMI),
                            Louvain.top = c(tab$Louvain_homogenity_top, 
                                            tab$Louvain_completeness_top, 
                                            tab$Louvain_NMI_top),
                            Louvain.middle = c(tab$Louvain_homogenity_middle, 
                                               tab$Louvain_completeness_middle,
                                               tab$Louvain_NMI_middle))




longform.x = cbind.data.frame(method = c(rep(rep('HCD', dim(tab)[1]),3),
                                         rep(rep('Louvain', dim(tab)[1]),3)),
                              metric = as.factor(c(rep('homogeneity', dim(tab)[1]),
                                                   rep('completeness', dim(tab)[1]),
                                                   rep('NMI', dim(tab)[1]))),
                              TOAL = as.factor(rep(rep(tab$TOAL_Loss, 3),2)),
                              network = as.factor(rep(rep(tab$Network_type, 3),2)),
                              input_graph = as.factor(rep(rep(tab$input_graph, 3),2)),
                              performance.top = c(c(tab$Top_homogeneity, 
                                                    tab$Top_completeness, 
                                                    tab$Top_NMI),
                                                  c(tab$Louvain_homogenity_top, 
                                                    tab$Louvain_completeness_top, 
                                                    tab$Louvain_NMI_top)),
                              performance.middle = c(c(tab$Middle_homogeneity, 
                                                       tab$Middle_Completeness, 
                                                       tab$Middle_NMI),
                                                     c(tab$Louvain_homogenity_middle, 
                                                       tab$Louvain_completeness_middle,
                                                       tab$Louvain_NMI_middle)))



# comparing HCD with louvain on graph type 
A = ggplot(data = longform.x, aes(x = method, y = performance.top, fill = metric))+
  geom_boxplot()+
  xlab('Method')+
  ylab('Performance Top Layer')+
  facet_wrap(~network)+
  theme_classic2()+
  theme(legend.position = 'top')

B = ggplot(data = longform.x, aes(x = method, y = performance.middle, fill = metric))+
  geom_boxplot()+
  xlab('Method')+
  ylab('Performance Middle Layer')+
  facet_wrap(~network)+
  theme_classic2()+
  theme(legend.position = 'top')


ggarrange(A, B, nrow = 2, common.legend = T)
```



```{r, echo=F, message=F, warning=F, fig.height=8, fig.width=10}


# comparing HCD with louvain on input graph 
A = ggplot(data = longform.x, aes(x = method, y = performance.top, fill = metric))+
  geom_boxplot()+
  xlab('Method')+
  ylab('Performance Top Layer')+
  facet_wrap(~input_graph)+
  theme_classic2()+
  theme(legend.position = 'top')

B = ggplot(data = longform.x, aes(x = method, y = performance.middle, fill = metric))+
  geom_boxplot()+
  xlab('Method')+
  ylab('Performance Middle Layer')+
  facet_wrap(~input_graph)+
  theme_classic2()+
  theme(legend.position = 'top')


ggarrange(A, B, nrow = 1, common.legend = T)

```




```{r, echo=F, message=F, warning=F, fig.height=8, fig.width=10, fig.cap='Small world graphs'}
longform2 = cbind.data.frame(method = as.factor(c(rep('HCD', dim(longform)[1]), 
                                        rep('Louvain', dim(longform)[1]))),
                             metric = as.factor(rep(c(rep('homogeneity', dim(tab)[1]),
                                         rep('completeness', dim(tab)[1]),
                                         rep('NMI', dim(tab)[1])), 2)),
                  network = as.factor(rep(tab$Network_type, 6)),
                  input_graph = as.factor(rep(tab$input_graph, 6)),
                  gamma = as.factor(rep(tab$Gamma, 6)),
                  resolution_top = as.factor(rep(tab$Resolution_top, 6)),
                  resolution_middle = as.factor(rep(tab$Resolution_middle, 6)),
                  stats.top = c(longform$Top.stats, longform$Louvain.top),
                  stats.mid = c(longform$Mid.stats, longform$Louvain.top))
                  
                  
                  
                  
#Top level comparison
#smallworld network
lf1.1 = subset(longform2, network == 'small world')
A =ggplot(data = lf1.1, aes(x = method, y = stats.top, 
                             fill = metric))+
  geom_boxplot()+
  xlab('Method')+
  ylab('Top Layer Performance')+
  facet_wrap(~input_graph)+
  theme_classic2()+
  theme(legend.position = 'top', axis.text.x = element_text(angle = 90))


B =ggplot(data = lf1.1, aes(x = method, y = stats.mid, 
                         fill = metric))+
  geom_boxplot()+
  xlab('Method')+
  ylab('Middle Layer Performance')+
  facet_wrap(~input_graph)+
  theme_classic2()+
  theme(legend.position = 'top', axis.text.x = element_text(angle = 90))

ggarrange(A, B, nrow = 1, common.legend = T)

```


```{r, echo=F, message=F, warning=F, fig.height=8, fig.width=10, fig.cap='Scale free graphs'}
#scale free
lf1.2 = subset(longform2, network == 'scale free')
A = ggplot(data = lf1.2, aes(x = method, y = stats.top, 
                             fill = metric))+
  geom_boxplot()+
  xlab('Method')+
  ylab('Top Layer Performance')+
  facet_wrap(~input_graph)+
  theme_classic2()+
  theme(legend.position = 'top', axis.text.x = element_text(angle = 90))

B = ggplot(data = lf1.2, aes(x = method, y = stats.mid, 
                         fill = metric))+
  geom_boxplot()+
  xlab('Method')+
  ylab('Middle Layer Performance')+
  facet_wrap(~input_graph)+
  theme_classic2()+
  theme(legend.position = 'top', axis.text.x = element_text(angle = 90))

ggarrange(A, B, nrow = 1, common.legend = T)
```


```{r, echo=F, message=F, warning=F, fig.height=8, fig.width=10, fig.cap='random graphs'}
#random graph
lf1.3 = subset(longform2, network == 'random graph')
A = ggplot(data = lf1.3, aes(x = method, y = stats.top, 
                             fill = metric))+
  geom_boxplot()+
  xlab('Method')+
  ylab('Top Layer Performance')+
  facet_wrap(~input_graph)+
  theme_classic2()+
  theme(legend.position = 'top', axis.text.x = element_text(angle = 90))

B = ggplot(data = lf1.3, aes(x = method, y = stats.mid, 
                         fill = metric))+
  geom_boxplot()+
  xlab('Method')+
  ylab('Middle Layer Performance')+
  facet_wrap(~input_graph)+
  theme_classic2()+
  theme(legend.position = 'top', axis.text.x = element_text(angle = 90))

ggarrange(A, B, nrow = 1, common.legend = T)
```








\newpage 
\section*{Tables}

```{r, echo=F, message=F, warning=F}

net_stats = read.csv('C:/Users/Bruin/Documents/GitHub/HGRN_repo/Simulated Hierarchies/DATA/Toy_examples/Intermediate_examples/intermediate_examples_network_statistics.csv')[,-1]

net_stats$modularity_top = round(net_stats$modularity_top, 3)
net_stats$avg_node_degree_top = round(net_stats$avg_node_degree_top, 3)
net_stats$avg_connect_within_top = round(net_stats$avg_connect_within_top, 3)
net_stats$avg_connect_between_top = round(net_stats$avg_connect_between_top, 3)
net_stats$modularity_middle = round(net_stats$modularity_middle, 3)
net_stats$avg_node_degree_middle = round(net_stats$avg_node_degree_middle, 3)
net_stats$avg_connect_within_middle = round(net_stats$avg_connect_within_middle, 3)
net_stats$avg_connect_between_middle = round(net_stats$avg_connect_between_middle, 3)

df = rbind.data.frame(c('Subgraph type','Connection type','Layers','Standard deviation','Nodes per layer','Edges per layer','Subgraph probability','Sample size', 'Modularity (top)', 'Average node degree top', 'Avg connections within top communities','Avg. connections between top communities', 'Modularity (middle)', 'Average node degree middle', 'Avg connections within middle communities', 'Avg connections between middle communities'), net_stats)

kable(t(df), format = 'latex', digits = 2, row.names = F, col.names = c('Value', paste0('Network', c(1:6))),
                                                                                    booktabs = T,
      caption = 'Summary statistics for intermediate difficulty simulated networks.')%>%kable_styling(latex_options = c("striped", "hold_postion"),font_size = 10)%>%kable_styling(font_size = 10)%>%column_spec(1, '8em')
```

\clearpage
\newpage
```{r, echo=F, message=F, warning=F}

grid = expand.grid(`Input Graph` = unique(tab$input_graph),
                   `Graph Recon. Loss` = c('1 = on', '0 = off'),
                   `Attr. Recon. Loss` = c('False (on)', 'True (off)'),
                   `Modularity Weigth` = c('1 = on', '0 = off'),
                   `Clust. Weight` = c('1 (middle), 1 (top)', '0.1 (middle), 1e-4 (top)'))

df = cbind.data.frame(`Scenario` = c(1:dim(grid)[1]), grid)
kable(df, format = 'markdown', digits = 2, row.names = F, booktabs = T,
      caption = 'Simulation settings for intermediate difficulty networks. Each row represents a single simulation scenario applied to all 6 simulated networks given in Table 1')
```


\bibliographystyle{unsrt}
	\bibliography{C:/Users/Bruin/Desktop/Research Assistantship/Thesis Proposal Defense/proposal_references.bib}

\newpage
\section*{References}
